{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}

### 2016-09-27 ident_manu.Rnw
added better tables and figures for initial and structural analyses

  }) %>% 
  do.call('rbind', .) %>%
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank)
    ) %>%
  group_by(stts) %>% 
  nest %>% 
  unnest %>% 
  select(stts, Category, Parameter, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\footnotesize{', Parameter, '}')
    ) %>% 
  spread(stts, coll, fill = '-') %>% 
  group_by(Parameter) %>% 
  arrange(Category, Rank)

@@

latex( 
  totab[, -c(1, 2)],
  file = '',
  rowlabel = 'Parameter',
  caption = cap_val,
  caption.loc = 'top',
  size = 'footnotesize',
  label = 'tab:heurist1',
  rowname = parms,
  rgroup = unique(cats),
  n.rgroup = as.numeric(table(cats)),
  colheads = colnms,
  rgroupTexCmd = NULL
  )
@

@@

# format tocal_all for identifiability table
totab <- lapply(tocal_all, function(x) x$top) %>% 
  reshape2::melt(., id.var = names(.[[1]])) %>%
  mutate(Parameter = as.character(Parameter)) %>% 
  mutate(
    stts = factor(L1, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = cats_lev, labels = cats_lab), 
    Parameter = par_txt(Parameter, frm = 'tex')
  ) %>% 
  filter(coll < 15) %>% 
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank, '\\textsuperscript{', Category, '}')
    ) %>%
  select(stts, Selected, Parameter, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\footnotesize{', Parameter, '}')
    ) %>% 
  arrange(stts, Selected)

@@

latex( 
  totab[, -c(1, 2)],
  file = '',
  rowlabel = 'Selections by state variable',
  caption = cap_val,
  caption.loc = 'top',
  size = 'footnotesize',
  label = 'tab:heurist2',
  rowname = sels,
  rgroup = unique(stts),
  n.rgroup = as.numeric(table(stts)),
  colheads = colnms,
  rgroupTexCmd = NULL
  )
@

@@

# format tocal_all for identifiability table
totab <- lapply(tocal_all, function(x) x$cattop) %>% 
  reshape2::melt(., id.var = names(.[[1]])) %>%
  mutate(Parameter = as.character(Parameter)) %>% 
  mutate(
    stts = factor(L1, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = cats_lev, labels = cats_lab), 
    Parameter = par_txt(Parameter, frm = 'tex')
  ) %>% 
  filter(coll < 15) %>% 
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank, '\\textsuperscript{', Category, '}')
    ) %>%
  select(stts, Selected, Parameter, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\footnotesize{', Parameter, '}')
    ) %>% 
  arrange(stts, Selected)

@@

latex( 
  totab[, -c(1, 2)],
  file = '',
  rowlabel = 'Selections by state variable',
  caption = cap_val,
  caption.loc = 'top',
  size = 'footnotesize',
  label = 'tab:heurist3',
  rowname = sels,
  rgroup = unique(stts),
  n.rgroup = as.numeric(table(stts)),
  colheads = colnms,
  rgroupTexCmd = NULL
  )
@

@@

<<sensalltile, eval = T, echo = F, fig.height = 7, fig.width = 6, out.width = '\\textwidth', fig.cap = 'Sensitivity values (L1, \\cref{l1}) for local analyses of all state variables. Parameters are grouped by category: optics, organic matter, phytoplankton, zooplankton, temperature, and zoplankton.  See \\cref{tab:dosens} for L1 values for \\ac{do} and \\cref{tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens} for the other state variables.'>>=
data(sens_ests_cat_all)

@@

ggplot(toplo, aes(x = Selected, y = coll)) + 
  geom_line() + 
  geom_text_repel(aes(label = Parameter), size = 3,
    parse = TRUE,
    point.padding = unit(0, "lines"),
    box.padding = unit(1, "lines"), 
    force = 3
    ) +
  geom_point(aes(fill = Category), shape = 21, size = 4) +
  geom_hline(yintercept = 15, linetype = 'longdash', colour = 'darkgrey') + 
  geom_hline(yintercept = 10, linetype = 'dotted', colour = 'darkgrey') + 
  facet_wrap(~L1, scales= 'free', ncol = 4) + 
  scale_y_continuous('Identifiability') + 
  scale_x_continuous(
    'Parameters selected by sensitivity',
    breaks = toplo$Selected, 
    expand = c(.1, .1)
    ) + 
  cols +
  pthm +
  theme(legend.position = 'top')
@
\clearpage

@@

% effects of structural uncertainty
<<streval, fig.cap = 'Effects of model structure on parameter sensitivity.', fig.height = 8, fig.width = 8, out.width = '\\textwidth', eval = F, echo = F, warning = F, message = F>>=
# structural
data(struct_eval)

swtc_levs <- paste0('Which_quota', 1:length(struct_eval))
swtc_labs <- paste0('Which quota: ', c('Droop', 'Nyholm', 'Flynn'))

toplo_all <- lapply(struct_eval, function(x) {

  tmp <- reshape2::melt(x[[2]], id.var = c('Parameter', 'error', 'Category'))
  swtc <- paste0(names(x[[1]]), x[[1]])

@@

  toplo <- mutate(tmp,
      L1 = factor(L1, levels = rev(stts_lev), labels = rev(stts_lab)),
      Parameter = as.character(par_txt(Parameter, frm = 'exp')),
      swtc = as.character(swtc)
    ) %>% 
    rename(Sensitivity = error)

@@

  })
toplo_all <- do.call('rbind', toplo_all)
toplo_all <- mutate(toplo_all, swtc = factor(swtc, levels = swtc_levs, labels = swtc_labs))

@@

pstr <- ggplot(toplo_all, aes(y = Parameter, x = L1, fill = Sensitivity)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Parameter', 
    expand = c(0, 0),
    breaks = toplo_all$Parameter, 
    labels = parse(text = toplo_all$Parameter)
    ) + 
  scale_x_discrete('State variable', expand = c(0, 0)) +
  scale_fill_distiller(palette = "Spectral", trans = 'log10',  na.value = 'grey95') +
  facet_grid(Category~swtc, scales = 'free_y', space = 'free') + 
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10))

print(pstr)
@

% effects of initial conditions
<<inieval, fig.cap = 'Effects of initial conditions on parameter sensitivity.', fig.height = 8, fig.width = 8, out.width = '\\textwidth', eval = T, echo = F, warning = F, message = F>>=
# initial conditions
data(ini_eval)

ini_levs <- paste('Initial:', c('def', 'apr', 'sep'))
ini_labs <- paste('Initial:', c('Default', 'April', 'September'))

toplo_all <- sapply(1:length(ini_eval), function(x) {

  tmp <- reshape2::melt(ini_eval[[x]][[2]], id.var = c('Parameter', 'error', 'Category'))
  ini <- names(ini_eval)[x]

  toplo <- mutate(tmp,
      L1 = factor(L1, levels = rev(stts_lev), labels = rev(stts_lab)),
      Parameter = as.character(par_txt(Parameter, frm = 'exp')), 
      ini = paste('Initial:', ini)
    ) %>% 
    rename(Sensitivity = error)

  }, simplify = F)
toplo_all <- do.call('rbind', toplo_all)
toplo_all <- mutate(toplo_all, ini = factor(ini, levels = ini_levs, labels = ini_labs))

pini <- ggplot(toplo_all, aes(y = Parameter, x = L1, fill = Sensitivity)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Parameter', 
    expand = c(0, 0),
    breaks = toplo_all$Parameter, 
    labels = parse(text = toplo_all$Parameter)
    ) + 
  scale_x_discrete('State variable', expand = c(0, 0)) +
  scale_fill_distiller(palette = "Spectral", trans = 'log10',  na.value = 'grey95') +
  facet_grid(Category~ini, scales = 'free_y', space = 'free') + 
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10))

print(pini)
@

>>>>

  }) %>% 
  do.call('rbind', .) %>%
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank)
    ) %>%
  group_by(stts) %>% 
  nest %>% 
  unnest %>% 
  select(stts, Category, Parameter, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\scriptsize{', Parameter, '}')
    ) %>% 
  spread(stts, coll, fill = '-') %>% 
  group_by(Parameter) %>% 
  arrange(Category, Rank)

@@

latex( 
  totab[, -c(1, 2)],
  file = '',
  rowlabel = 'Parameter',
  caption = cap_val,
  caption.loc = 'top',
  size = 'scriptsize',
  label = 'tab:heurist1',
  rowname = parms,
  rgroup = unique(cats),
  n.rgroup = as.numeric(table(cats)),
  colheads = colnms,
  rgroupTexCmd = NULL
  )
@

@@

# format tocal_all for identifiability table
totab <- lapply(tocal_all, function(x) x$top) %>% 
  reshape2::melt(., id.var = names(.[[1]])) %>%
  mutate(Parameter = as.character(Parameter)) %>% 
  mutate(
    stts = factor(L1, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = cats_lev, labels = cats_lab), 
    Parameter = par_txt(Parameter, frm = 'tex')
  ) %>% 
  filter(coll < 15) %>% 
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank, '\\textsuperscript{', Category, '}')
    ) %>%
  select(stts, Selected, Parameter, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\scriptsize{', Parameter, '}')
    ) %>% 
  arrange(stts, Selected)

@@

latex( 
  totab[, -c(1, 2)],
  file = '',
  rowlabel = 'Selections by state variable',
  caption = cap_val,
  caption.loc = 'top',
  size = 'scriptsize',
  label = 'tab:heurist2',
  rowname = sels,
  rgroup = unique(stts),
  n.rgroup = as.numeric(table(stts)),
  colheads = colnms,
  rgroupTexCmd = NULL
  )
@

@@

# format tocal_all for identifiability table
totab <- lapply(tocal_all, function(x) x$cattop) %>% 
  reshape2::melt(., id.var = names(.[[1]])) %>%
  mutate(Parameter = as.character(Parameter)) %>% 
  mutate(
    stts = factor(L1, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = cats_lev, labels = cats_lab), 
    Parameter = par_txt(Parameter, frm = 'tex')
  ) %>% 
  filter(coll < 15) %>% 
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank, '\\textsubscript{', Category, '}')
    ) %>%
  select(stts, Selected, Parameter, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\scriptsize{', Parameter, '}')
    ) %>% 
  arrange(stts, Selected)

@@

latex( 
  totab[, -c(1, 2)],
  file = '',
  rowlabel = 'Selections by state variable',
  caption = cap_val,
  caption.loc = 'top',
  size = 'scriptsize',
  label = 'tab:heurist3',
  rowname = sels,
  rgroup = unique(stts),
  n.rgroup = as.numeric(table(stts)),
  colheads = colnms,
  rgroupTexCmd = NULL
  )
@

% change in sensitivity for different initial, structural conditions
<<inistrchg, eval = T, cache = F>>=
# initial conditions
data(ini_eval)
data(struct_eval)

to_eval <- lapply(ini_eval, function(x) x[[2]]) %>% 
  c(., struct = list(struct_eval))

eval_levs <- c('def', 'apr', 'sep', 'struct')
eval_labs <- c('Default', 'Initial: April', 'Initial: September', 'Structure: complex')

totab_all <- sapply(1:length(to_eval), function(x) {

  tmp <- reshape2::melt(to_eval[[x]], id.var = c('Parameter', 'error', 'Category'))
  nms <- names(to_eval)[x]

  toplo <- mutate(tmp,
      L1 = factor(L1, levels = rev(stts_lev), labels = rev(stts_lab)),
      Parameter = as.character(par_txt(Parameter, frm = 'exp')), 
      nms = nms
    ) %>% 
    rename(Sensitivity = error)

  }, simplify = F)
totab_all <- do.call('rbind', totab_all)
totab_all <- mutate(totab_all, nms = factor(nms, levels = eval_levs, labels = eval_labs))

# summarize
totab_all <- spread(totab_all, nms, Sensitivity) %>% 
  group_by(Category, L1) %>% 
  summarise(
    mean_def = mean(Default, na.rm = TRUE),
    iniapr_dif = mean_def - mean(`Initial: April`, na.rm = TRUE),
    iniapr_sgn = sign(iniapr_dif),
    iniapr_dif = sapply(iniapr_dif, scinot),
    iniapr_dif = ifelse(iniapr_sgn == -1, iniapr_dif, gsub('\\$$', '}$', gsub('^\\$', '$\\\\bm{', iniapr_dif))),
    inisep_dif = mean_def - mean(`Initial: September`, na.rm = TRUE),
    inisep_sgn = sign(inisep_dif),
    inisep_dif = sapply(inisep_dif, scinot),
    inisep_dif = ifelse(inisep_sgn == -1, inisep_dif, gsub('\\$$', '}$', gsub('^\\$', '$\\\\bm{', inisep_dif))),
    strcom_dif = mean_def - mean(`Structure: complex`, na.rm = TRUE),
    strcom_sgn = sign(strcom_dif),
    strcom_dif = sapply(strcom_dif, scinot),
    strcom_dif = ifelse(strcom_sgn == -1, strcom_dif, gsub('\\$$', '}$', gsub('^\\$', '$\\\\bm{', strcom_dif))),
    mean_def = sapply(mean_def, scinot)
  ) %>% 
  select(-iniapr_sgn, -inisep_sgn, -strcom_sgn)

# final table formatting
cats <- as.character(totab_all$Category)
stts<- totab_all$L1
cap_val<- 'Changes in mean sensitivity of state variables by parameter categories to different initial conditions and changes in structural components of the model.  Changes show the difference in average sensitivity from results using the default model setup. Sensitivities are based on average $L1$ (see \\cref{l1}) values of the state variables to changes in parameters in each parameter category. Increases in average sensitivity are in bold. Medians and ranges of sensitivity values for the different conditions are shown in \\cref{fig:inistrsumm}.'
colnms <- c('Default sensitivity', 'Initial: April', 'Initial: September', 'Structure: complex')

latex( 
  totab_all[, -c(1, 2)],
  file = '',
  rowlabel = 'State variables by parameter category',
  caption = cap_val,
  caption.loc = 'top',
  size = 'scriptsize',
  label = 'tab:inistrchg',
  rowname = stts,
  rgroup = unique(cats),
  n.rgroup = as.numeric(table(cats)),
  colheads = colnms,
  rgroupTexCmd = NULL
  )

@@

<<sensalltile, eval = T, echo = F, fig.height = 6, fig.width = 3.5, out.width = '0.7\\textwidth', fig.cap = 'Sensitivity values (L1, \\cref{l1}) for local analyses of all state variables. Parameters are grouped by category: optics, organic matter, phytoplankton, zooplankton, temperature, and zoplankton.  See \\cref{tab:dosens} for L1 values for \\ac{do} and \\cref{tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens} for the other state variables.'>>=
data(sens_ests_cat_all)

@@

ggplot(toplo, aes(x = Selected, y = coll)) + 
  geom_line() + 
  geom_text_repel(aes(label = Parameter), size = 3,
    parse = TRUE,
    point.padding = unit(0, "lines"),
    box.padding = unit(1, "lines"), 
    force = 3
    ) +
  geom_point(aes(fill = Category), shape = 21, size = 4) +
  geom_hline(yintercept = 15, linetype = 'longdash', colour = 'darkgrey') + 
  geom_hline(yintercept = 10, linetype = 'dotted', colour = 'darkgrey') + 
  facet_wrap(~L1, scales= 'free', ncol = 4) + 
  scale_y_continuous('Identifiability') + 
  scale_x_continuous(
    'Parameters selected by sensitivity',
    breaks = toplo$Selected, 
    expand = c(.1, .1)
    ) + 
  cols +
  pthm +
  theme(legend.position = 'top')
@

% summary of initial/structural changes on parameter sensitivity
<<inistrsumm, eval = T, fig.cap = 'Summaries of changes in parameter sensitivity values (L1, \\cref{l1}) from the default model conditions.  Sensitivity was re-evaluated using initial conditions as seasonal means for April and September, and added structural complexity.  Sensitivity is summarized as the minimum, median, and maximum for each state variable separated by parameter categories. Changes in average sensitivity from the default model conditions are in \\cref{tab:inistrchg}.', fig.height = 7, fig.width = 7, echo = F, cache = TRUE, out.width = '\\textwidth'>>=

# initial conditions
data(ini_eval)
data(struct_eval)

to_eval <- lapply(ini_eval, function(x) x[[2]]) %>% 
  c(., struct = list(struct_eval))

eval_levs <- c('def', 'apr', 'sep', 'struct')
eval_labs <- c('Default', 'Initial: April', 'Initial: September', 'Structure: complex')

toplo_all <- sapply(1:length(to_eval), function(x) {

  tmp <- reshape2::melt(to_eval[[x]], id.var = c('Parameter', 'error', 'Category'))
  nms <- names(to_eval)[x]

  toplo <- mutate(tmp,
      L1 = factor(L1, levels = rev(stts_lev), labels = rev(stts_lab)),
      Parameter = as.character(par_txt(Parameter, frm = 'exp')), 
      nms = nms
    ) %>% 
    rename(Sensitivity = error)

  }, simplify = F)
toplo_all <- do.call('rbind', toplo_all)
toplo_all <- mutate(toplo_all, nms = factor(nms, levels = eval_levs, labels = eval_labs))

# summarize
toplo_all <- group_by(toplo_all, Category, L1, nms) %>% 
  summarise(
    med_val = median(Sensitivity, na.rm = TRUE),
    ave_val = mean(Sensitivity, na.rm = TRUE), 
    min_val = quantile(Sensitivity, 0, na.rm = TRUE), 
    max_val = quantile(Sensitivity, 1, na.rm = TRUE)
  )

pbase <- theme_minimal() + 
  theme(
    legend.position = 'top',
    axis.text.x = element_text(size = 8), 
    axis.line.x = element_line(),
    axis.line.y = element_line(), 
    # panel.grid.minor = element_blank(), 
    # panel.grid.major = element_blank(),
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm")
    )

mycols <- RColorBrewer::brewer.pal(5, 'Spectral')
names(mycols) <- cats
fils<- scale_fill_manual(name = 'Category', values = mycols)
cols <- scale_colour_manual(name = 'Category', values = mycols)

ddge_wd <- 0.7
ggplot(toplo_all, aes(x = L1)) + 
  geom_errorbar(aes(ymin = min_val, ymax = max_val, group = Category), colour = 'black', width = 0, position = position_dodge(width = ddge_wd)) +
  geom_point(aes(y = med_val, fill = Category), colour = 'black', pch = 21, position = position_dodge(width = ddge_wd), size = 3) + 
  scale_y_continuous('Sensitivity', trans = 'log10') + 
  scale_x_discrete('State variable') + 
  facet_wrap( ~ nms, ncol = 1) +
  cols +
  fils + 
  pbase
@

@@

% effects of changing initial and structural conditions
<<inistreval, fig.cap = 'Effects of initial and structural conditions on parameter sensitivity.', fig.height = 7, fig.width = 8, out.width = '\\textwidth', eval = T, echo = F, warning = F, message = F>>=
# initial conditions
data(ini_eval)
data(struct_eval)

to_eval <- lapply(ini_eval, function(x) x[[2]]) %>% 
  c(., struct = list(struct_eval))

eval_levs <- c('def', 'apr', 'sep', 'struct')
eval_labs <- c('Default', 'Initial: April', 'Initial: September', 'Structure: complex')

toplo_all <- sapply(1:length(to_eval), function(x) {

  tmp <- reshape2::melt(to_eval[[x]], id.var = c('Parameter', 'error', 'Category'))
  nms <- names(to_eval)[x]

@@

  toplo <- mutate(tmp,
      L1 = factor(L1, levels = rev(stts_lev), labels = rev(stts_lab)),
      Parameter = as.character(par_txt(Parameter, frm = 'exp')), 
      nms = nms
    ) %>% 
    rename(Sensitivity = error)

@@

  }, simplify = F)
toplo_all <- do.call('rbind', toplo_all)
toplo_all <- mutate(toplo_all, nms = factor(nms, levels = eval_levs, labels = eval_labs))

@@

p <- ggplot(toplo_all, aes(y = Parameter, x = L1, fill = Sensitivity)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('Parameter', 
    expand = c(0, 0),
    breaks = toplo_all$Parameter, 
    labels = parse(text = toplo_all$Parameter)
    ) + 
  scale_x_discrete('State variable', expand = c(0, 0)) +
  scale_fill_distiller(palette = "Spectral", trans = 'log10',  na.value = 'grey95') +
  facet_grid(Category~nms, scales = 'free_y', space = 'free') + 
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10))

print(p)
@
